\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\author{Kees Mulder}
\title{A Circular Outcome Bayesian GLM}
\begin{document}

\maketitle

\section{Introduction}

\begin{itemize}
\item Previous work provides a distribution (Mardia), and a GLM-type model (Lagona). Earlier work was very limited (Gill, Hangartner).
\item Full bayesian analysis requires a sampler, priors, posteriors. Some properties of MCMC sampling for the von Mises are discussed in my previous work.
\item Here, we will develop a flexible method to analyse circular outcome data, which can also be used in ANCOVA type situations. 
\end{itemize}

\section{The model}

\begin{itemize}
\item Multivariate von Mises distribution
\item Other multivariate circular distributions?
\item Notation
\item Link function
\item Likelihood
\item Prior
\item Posterior
\end{itemize}

\section{Computational issues}

\begin{itemize}
\item Shape of the marginal likelihood for some common cases
\item Selection of starting values
\item Non-interpretability of beta's
\item Solutions
\end{itemize}

\section{Simulation study}

\begin{itemize}
\item Cases
\item Results
\item Figures
\end{itemize}

\section{Application}

\begin{itemize}
\item 1-3 example datasets, analyzed with CCI's etc. 
\item At least one highly dimensional dataset that would fail under Gill \& Hangartner.  
\end{itemize}


\section{Discussion}

\begin{itemize}
\item This model is great.
\item This model is far superior to the projected approach, because of reasons. Or maybe worse. 
\item Can be extended with association matrix, multilevel structures (Lagona).
\end{itemize}

\end{document}